Edge sign prediction in general graphs and Correlation Clustering

### 2017-09-04 troll/experiments.tex
a pass on the rest of the troll main text

\section{Experimental Analysis}\label{s:exp}
We now evaluate our edge sign classification methods on representative real-world datasets of varying density and label regularity,
% within the batch settings,
showing that our methods compete well against existing approaches in terms of both predictive and computational performance. We are especially interested in small training set regimes, and have restricted our comparison to the batch learning scenario since all competing methods we are aware of have been developed in that setting only.

{\bf Datasets.} We considered five real-world classification datasets. The first three are directed \ssn{} widely used as benchmarks for this task~(e.g.,\cite{Leskovec2010,shahriari2014ranking,wu2016troll}): In \wik{}, there is an edge from user $i$ to user $j$ if $j$ applies for an admin position and $i$ votes for or against that promotion. In \sla{}, a news sharing and commenting website, member $i$ can tag other members $j$ as friends or foes. Finally, in \epi{}, an online shopping website, user $j$ reviews products and, based on these reviews, another user $i$ can display whether he considers $j$ to be reliable or not. In addition to these three datasets, we considered two other \ssn{} where the signs are inferred automatically.
%rather than given explicitly by the users.
In \kiw{}~\cite{wikiedits11}, an edge from Wikipedia user $i$ to user $j$ indicates whether they edited the same article in a constructive manner or not.\footnote
{
This is the \href{http://konect.uni-koblenz.de/networks/wikisigned-k2}{KONECT version of
the \enquote{Wikisigned} dataset}, from which we removed self-loops.
}
Finally, in the \aut{}~\cite{kumar2016structure} network, an author $i$ cites another author $j$ by either endorsing or criticizing $j$'s work. The edge sign is derived by classifying the citation sentiment with a simple, yet powerful, keyword-based technique using a list of positive and negative words. See \cite{kumar2016structure} for more details.\footnote{
We again removed self-loops and merged multi-edges which are all of the same sign.
}

\begin{table*}[bt]
  \centering
  \small
  \caption{Dataset properties. The 5th column gives the fraction of positive labels. The last two columns provide two different measures of label regularity ---see main text.\label{tab:dataset}}
  \begin{tabular}{lrrrrrrrr}
    \toprule
    Dataset &       $|V|$ &       $|E|$ &$\frac{|E|}{|V|}$ & $\frac{|E^+|}{|E|}$ & $\frac{\Psi^2_{G''}(Y)}{|E|}$ & $\frac{\Psi_G(Y)}{|E|}$ \\ % & twin edges & twin disagreement \\
    \midrule
    \aut{} &   \np{4831} &  \np{39452} & 8.1  &              72.33\% &                    .076 &                       .191 \\ %&      5.07\% &            27.10\% \\
    \wik{} &   \np{7114} & \np{103108} & 14.5 &              78.79\% &                    .063 &                       .142 \\ %&      5.56\% &            10.04\% \\
    \sla{} &  \np{82140} & \np{549202} & 6.7  &              77.40\% &                    .059 &                       .143 \\ %&     17.74\% &             4.00\% \\
    \kiw{} & \np{138587} & \np{740106} & 5.3  &              87.89\% &                    .034 &                       .086 \\ %&      6.55\% &             2.09\% \\
    \epi{} & \np{131580} & \np{840799} & 6.4  &              85.29\% &                    .031 &                       .074 \\ %&     30.83\% &            14.64\% \\
    \bottomrule
  \end{tabular}
\end{table*}

\autoref{tab:dataset} summarizes statistics for these datasets. We note that most edge labels are positive. Hence, test set accuracy is not an appropriate measure of prediction performance. We instead evaluated our performance using the so-called Matthews Correlation Coefficient (MCC)~(e.g., \cite{MCC00}), defined as
%\footnote
%{
%For completeness, we also report accuracy results in the supplementary material.
%}  
\[
	\mathrm{MCC} = \frac{tp\times tn-fp\times fn} {\sqrt{ (tp + fp) ( tp + fn ) ( tn + fp ) ( tn + fn ) } }\,.
			% = \pm \sqrt{\frac{\chi^2}{n}}
\]
MCC combines all the four quantities found in a binary confusion matrix ($t$rue $p$ositive, $t$rue $n$egative, $f$alse $p$ositive and $f$alse $n$egative) into a single metric which ranges from $-1$ (when all predictions are incorrect) to $+1$ (when all predictions are correct).% through $0$ (when predictions are made uniformly at random).

Although the semantics of the edge signs is not the same across these networks, we can see from \autoref{tab:dataset} that our generative model essentially fits all of them. Specifically, the last two columns of the table report the rate of label (ir)regularity, as measured by $\Psi^2_{G''}(Y)/|E|$ (second-last column) and $\Psi_{G}(Y)/|E|$ (last column), where 
\[
\Psi^2_{G''}(Y) = \min_{(\bp,\bq)} \left(f_{E_0}(\bp,\bq) + f_{E\setminus E_0}(\bp,\bq)\right)\,,
\]
%$\Psi^2_{G''}(Y) = \min_{(\bp,\bq)} f_{E}(\bp,\bq)$, 
%
$f_{E_0}$ and $f_{E\setminus E_0}$ being the quadratic criterions of Section~\ref{ss:passive}, viewed as functions of both $(\bp,\bq)$, and $y_{i,j}$,
% when all labels are known,
\iffalse
*****************
\[
\Psi_{G''}(Y) = \min_{\bp,\bq\in[0,1]^{|V|}}
\sum_{(i,j) \in E} \left(\frac{1+y_{i,j}}{2} - \frac{p_i+q_j}{2}\right)^2\,,
\]
i.e., the quadratic criterion $f_{E}(\bp,\bq)$ of Section~\ref{ss:passive} when all labels are known,
*****************
\fi
%essentially hinges upon, 
and $\Psi_{G}(Y)$ is the label regularity measure adopted in the online setting, as defined in Section~\ref{s:prel}. 
%The two measures have been crafted here so as to be on the same scale, hence they can be contrasted to each other on the same dataset. 
It is reasonable to expect that higher label irregularity corresponds to lower prediction performance. This trend is in fact confirmed by our experimental findings: whereas \epi{} tends to be easy, \aut{} tends to be hard, and this holds for all algorithms we tested, even if they do not explicitly comply with our inductive bias principles. Moreover, $\Psi^2_{G''}(Y)/|E|$ tends to be proportional to $\Psi_{G}(Y)/|E|$ across datasets, hence confirming the anticipated connection between the two regularity measures.

%
%Another notable feature of our datasets is the relative regularity of their
%labeling with respect to our two complexity measures defined in
%\autoref{s:analysis}
%(here  where the minimum
%is found using the L-BFGS-B algorithm)
%(especially in \epi{} case, which is confirmed by higher performance).
%

@@

{\bf Algorithms and parameter tuning.} 
We compared the following algorithms:
%
%\begin{itemize}%[nosep,leftmargin=2em]
%
%\item

{\bf 1.} The label propagation algorithm of Section~\ref{ss:passive} (referred to as \uslpropGsec{}). The actual binarizing threshold was set by cross-validation on the training set.
%
%\item

{\bf 2.} The algorithm analyzed at the beginning of Section~\ref{s:algbatch}, which we call \usrule{} (Bayes Learning Classifier based on {\em tr}ollness and {\em un}trustworthiness). After computing $\htr(i)$ and $\hun(i)$ on training set $E_0$ for all $i \in V$ (or setting those values to $\frac{1}{2}$ in case there is no outgoing or incoming edges for some node), we use Eq.~\eqref{eq:predictor} and estimate $\tau$ on $E_0$.
%
%\item

{\bf 3.} A logistic regression model where each edge $(i,j)$ is associated with the features $[1-\htr(i), 1-\hun(j)]$ computed again on $E_0$ (we call this method \uslogregp{}). Best binary thresholding is again computed on $E_0$. Experimenting with this logistic model serves to support the claim we made in the introduction that our generative model in Section~\ref{s:gen} is a good fit for the data.

%
%\item 
{\bf 4.} The solution obtained by directly solving the unregularized problem (\ref{e:quadratic}) through a fast constrained minimization algorithm (referred to as \qoptim{}). Again, the actual binarizing threshold was set by cross-validation on the training set.\footnote
{
We have also tried to minimize (\ref{e:quadratic}) by removing the $[-1,+1]$ constraints, but got similar MCC results
as the ones we report for \qoptim{}
}
%

{\bf 5.} The matrix completion method from~\cite{LowRankCompletion14} based on \complowrank{} matrix factorization. Since the authors showed their method to be robust to the choice of the rank parameter $k$, we picked $k=7$ in our experiments.
%
%\item

{\bf 6.} A logistic regression model built on \comptriads{} features derived from status theory~\cite{Leskovec2010}.
%
%\item

{\bf 7.} The PageRank-inspired algorithm from \cite{wu2016troll}, where a recursive notion of trollness is computed by solving a suitable set of nonlinear equations through an iterative method, and then used to assign ranking scores to nodes, from which (un)trustworthiness features are finally extracted for each edge. We call this method \compranknodes{}. As for hyperparameter tuning ($\beta$ and $\lambda_1$ in~\cite{wu2016troll}), we closely followed the authors' suggestion of doing cross validation.
%
%\item

{\bf 8.} The last competitor is the logistic regression model whose features have been build according to \cite{Bayesian15}. We call this method \compbayesian{}.
%\end{itemize}

The above methods can be roughly divided into {\em local} and {\em global} methods. A local method hinges on building local predictive features, based on neighborhoods: \usrule{}, \uslogregp{}, \comptriads{}, and \compbayesian{} essentially fall into this category. The remaining methods 
%(\uslpropGsec{}, \complowrank{}, and \compranknodes{}) 
are global in that their features are designed to depend on global properties of the graph topology.

@@

Looking at undirected graphs with a global perspective,
\citet{LowRankCompletion14} consider the observed adjacency matrix $A$, made of
the edges in \trainset{}, as a noisy sampling of the adjacency matrix
$A^\star$ of an underlying complete graph satisfying the weakly balance
condition (that is with no triangle containing only one negative edge).
%TODO the previous sentence is too long
This condition implies the existence of a small number $k$ of node clusters
with positive links within clusters and negative links across clusters, which
in turn implies $\rank A^\star=k$.
By recovering a complete matrix $\tilde{A}$ which matches the non-zeros entries
of $A$, it is possible to predict the sign of $\eij \notin \trainset$
as $\yhat_{i,j} = \sgn(\tilde{A}_{i,j})$. Although the exact version of this
problem is NP-hard, authors assume that $k$ is an hyperparameter known
beforehand\footnote{The method is showed to be robust to the choice of its
value, thus we pick $k=7$.} and look for two matrices $W,H\in
\mathbb{R}^{k\times|V|}$ which minimise a sigmoidal reconstruction loss with
$A$, subject to a regularization term. We refer to this method as
\emph{\complowrank{}}.

The status theory heuristic posits that a positive link from $i$ to $j$ denotes
that user $i$ considers user $j$ as having a higher status or skill than
himself~\cite{Leskovec2010}. It implies that among the 16 possible triads of
three nodes with two directed signed edges among them, some must be more
represented than others. The \emph{\comptriads{}} method exploits this fact by
counting for each edge in the training set how frequently it is involved in
each of the 16 triad types. Based on these edge features and 7 degree features
of each edge endpoints, a Logistic Regression model is trained and used to
predict the sign of the test edges.

Inspired by the PageRank algorithm, \citet{wu2016troll} propose a recursive
definition of trollness based on the opinion of one node's neighbors weighted
by their own trollness, which allow to assign a trustworthiness $\pi(i)$ to
each node $i$ through a set of non linear equations solved by an iterative
method. This $\pi$ values are used to rank the nodes by computing their
Reputation and Optimism scores, thus providing four features for each edge,
which are in turn used to train a Logistic Regression model for the
classification task. We refer to this method as \emph{\compranknodes{}}.
Computing $\pi$ requires to choose two hyperparameters of the model, $\beta$
and $\lambda_1$ and the authors suggest to do it by holding out some training
data to perform cross validation but in practice we found little difference in
performance in our implementation while it consumes a lot of time.

The last competitors is also local. \citet{Bayesian15} note that  a node can
belong to one of the 16 types based of whether is number of its positive (resp.
negative) outgoing (resp. incoming) edges is zero or not.  The number of
unobserved incoming and outgoing edges of each node $i$ let us define its
16-dimensional vector $V_i$ containing the probability of transitioning to any
other state.  Then for each edge \eij we compute the outer product of $V_i$
with $V_j$ and also include additional features such as triads count and degree
information before training a Logistic Regression model. We refer to this
method as \emph{\compbayesian{}}.

***************************************************
\fi

\iffalse
****************************************************

@@

% \begin{longtable}{lrccc|cccc}
\caption{MCC with increasing training set size, with one standard deviation over 12 random sampling of $\trainset$. The last four columns refer to the methods we took from the literature. For the sake of readability, we multiplied all MCC values by 100. The best number in each row is highlighted in \textbf{\textcolor{brown}{bold brown}} and the second one in \textit{\textcolor{red}{italic red}}. If the difference is statistically significant ($p$-value of a paired Student's $t$-test less than $0.005$), the best score is underlined. The ``time" rows contain the time taken to train on a $15\%$ training set.\label{tab:all_mcc}}
\begin{tabular}{lrccc|cccc}
                                                  & $\frac{|\trainset{}|}{|E|}$ &                    \uslogregp{} &                       \usrule{} &                 \uslpropGsec{} &   \compranknodes{} &                 \compbayesian{} &     \complowrank{} &      \comptriads{} \\
\multirow{7}{*}{\rotatebox[origin=c]{90}{\aut{}}}
                                                  & $3\%$  &  $\vsecondSig{15.19} \pm 0.93$  &               $15.09 \pm 0.92$  &  $\vfirstSig{19.00} \pm 1.16$  &  $12.28 \pm 1.70$  &               $10.91 \pm 1.31$  &   $8.85 \pm 1.07$  &   $8.62 \pm 0.97$  \\
                                                  %& $5\%$  &  $\vsecondSig{19.54} \pm 0.48$  &               $19.53 \pm 0.43$  &  $\vfirstSig{24.11} \pm 0.68$  &  $16.65 \pm 0.59$  &               $14.53 \pm 0.85$  &  $12.07 \pm 0.88$  &  $11.47 \pm 1.23$  \\
                                                  %& $7\%$  &  $\vsecondSig{23.58} \pm 0.55$  &               $23.55 \pm 0.43$  &  $\vfirstSig{27.64} \pm 1.08$  &  $21.38 \pm 0.86$  &               $19.75 \pm 1.02$  &  $15.15 \pm 0.54$  &  $14.58 \pm 1.15$  \\
                                                  & $9\%$  &  $\vsecondSig{26.46} \pm 0.96$  &               $26.40 \pm 0.96$  &  $\vfirstSig{30.25} \pm 1.16$  &  $24.44 \pm 1.07$  &               $23.75 \pm 1.37$  &  $17.08 \pm 1.15$  &  $16.42 \pm 1.13$  \\
                                                  & $15\%$ &  $\vsecondSig{32.98} \pm 0.61$  &               $32.98 \pm 0.65$  &  $\vfirstSig{35.73} \pm 0.70$  &  $31.03 \pm 0.97$  &               $32.25 \pm 0.89$  &  $22.57 \pm 0.99$  &  $22.01 \pm 0.62$  \\
                                                  & $20\%$ &               $36.57 \pm 0.49$  &  $\vsecondSig{36.72} \pm 0.44$  &  $\vfirstSig{38.53} \pm 0.55$  &  $34.57 \pm 0.79$  &               $36.52 \pm 0.78$  &  $25.57 \pm 0.71$  &  $24.77 \pm 0.64$  \\
                                                  & $25\%$ &               $39.90 \pm 0.65$  &               $40.16 \pm 0.63$  &  $\vfirstSig{41.32} \pm 0.67$  &  $38.26 \pm 0.77$  &  $\vsecondSig{40.32} \pm 0.70$  &  $29.24 \pm 1.10$  &  $27.13 \pm 0.38$  \\
                                                  & time (ms)   &                           1.9 &                           0.4 &                          16.1 &              128 &                         5398 &               1894 &              4.6 \\
\multirow{7}{*}{\rotatebox[origin=c]{90}{\wik{}}} 
                                                  & $3\%$  &  $\vsecondSig{32.32} \pm 0.60$  &  $31.83 \pm 0.63$  &   $\vfirstSig{33.92} \pm 0.54$  &  $26.90 \pm 1.00$  &  $19.94 \pm 1.19$  &  $19.45 \pm 0.64$  &   $4.29 \pm 0.80$  \\
                                                  %& $5\%$  &  $\vsecondSig{38.62} \pm 0.63$  &  $37.82 \pm 0.66$  &   $\vfirstSig{39.44} \pm 0.44$  &  $33.47 \pm 1.12$  &  $26.68 \pm 0.72$  &  $24.18 \pm 1.04$  &  $11.36 \pm 1.32$  \\
                                                  %& $7\%$  &     $\vsecond{42.81} \pm 0.80$  &  $42.15 \pm 0.76$  &      $\vfirst{43.28} \pm 0.61$  &  $38.13 \pm 0.80$  &  $33.18 \pm 1.27$  &  $27.89 \pm 0.64$  &  $18.57 \pm 1.38$  \\
                                                  & $9\%$  &     $\vsecond{45.57} \pm 0.48$  &  $44.74 \pm 0.58$  &      $\vfirst{45.75} \pm 0.46$  &  $41.60 \pm 0.73$  &  $38.25 \pm 1.02$  &  $30.75 \pm 0.68$  &  $24.04 \pm 1.27$  \\
                                                  & $15\%$ &      $\vfirst{50.70} \pm 0.28$  &  $49.64 \pm 0.57$  &     $\vsecond{50.44} \pm 0.38$  &  $48.02 \pm 0.38$  &  $46.82 \pm 0.40$  &  $35.31 \pm 0.75$  &  $34.42 \pm 1.00$  \\
                                                  & $20\%$ &   $\vfirstSig{52.98} \pm 0.50$  &  $52.00 \pm 0.47$  &  $\vsecondSig{52.58} \pm 0.40$  &  $51.42 \pm 0.33$  &  $50.45 \pm 0.40$  &  $38.16 \pm 0.53$  &  $38.55 \pm 0.72$  \\
                                                  & $25\%$ &      $\vfirst{54.49} \pm 0.30$  &  $53.52 \pm 0.41$  &     $\vsecond{54.22} \pm 0.44$  &  $53.42 \pm 0.36$  &  $52.78 \pm 0.61$  &  $39.94 \pm 0.48$  &  $41.51 \pm 0.46$  \\
                                                  & time (ms)   &                           4.2 &              1.1 &                           35.3 &              210 &              14089 &               4858 &              10.6 \\
\multirow{7}{*}{\rotatebox[origin=c]{90}{\sla{}}} 
                                                  & $3\%$  &  $32.34 \pm 0.27$  &  $31.78 \pm 0.81$  &  $\vsecondSig{36.62} \pm 0.42$  &   $\vfirstSig{42.90} \pm 0.54$  &  $25.11 \pm 0.66$  &  $34.32 \pm 0.64$  &  $20.95 \pm 0.86$  \\
                                                  %& $5\%$  &  $36.93 \pm 0.32$  &  $36.25 \pm 0.63$  &  $\vsecondSig{40.72} \pm 0.24$  &   $\vfirstSig{45.47} \pm 0.37$  &  $29.50 \pm 0.17$  &  $36.95 \pm 0.44$  &  $28.26 \pm 1.06$  \\
                                                  %& $7\%$  &  $39.96 \pm 0.27$  &  $39.26 \pm 0.39$  &  $\vsecondSig{43.62} \pm 0.21$  &   $\vfirstSig{46.80} \pm 0.64$  &  $33.57 \pm 0.27$  &  $38.25 \pm 0.31$  &  $33.89 \pm 1.41$  \\
                                                  & $9\%$  &  $42.16 \pm 0.16$  &  $41.19 \pm 0.30$  &  $\vsecondSig{45.70} \pm 0.26$  &   $\vfirstSig{47.46} \pm 0.27$  &  $37.00 \pm 0.17$  &  $39.42 \pm 0.20$  &  $39.14 \pm 1.43$  \\
                                                  & $15\%$ &  $46.44 \pm 0.11$  &  $45.23 \pm 0.32$  &   $\vfirstSig{49.65} \pm 0.23$  &  $\vsecondSig{48.59} \pm 0.72$  &  $43.28 \pm 0.21$  &  $41.09 \pm 0.29$  &  $46.27 \pm 1.00$  \\
                                                  & $20\%$ &  $48.71 \pm 0.22$  &  $47.79 \pm 0.29$  &     $\vsecond{51.88} \pm 0.28$  &      $\vfirst{52.09} \pm 0.28$  &  $47.03 \pm 0.24$  &  $43.10 \pm 0.42$  &  $49.44 \pm 0.76$  \\
                                                  & $25\%$ &  $50.23 \pm 0.16$  &  $49.43 \pm 0.22$  &     $\vsecond{53.30} \pm 0.15$  &      $\vfirst{53.46} \pm 0.23$  &  $49.46 \pm 0.24$  &  $44.37 \pm 0.50$  &  $51.51 \pm 0.48$  \\
                                                  & time (ms)   &              21.2 &              5.7 &                           655 &                            1918 &              77042 &              56252 &              78 \\
\multirow{7}{*}{\rotatebox[origin=c]{90}{\epi{}}} 
                                                  & $3\%$  &  $43.51 \pm 0.35$  &  $41.39 \pm 1.11$  &     $\vsecond{51.47} \pm 0.28$  &     $\vfirst{52.04} \pm 0.52$  &  $31.00 \pm 0.66$  &  $36.84 \pm 0.43$  &  $34.42 \pm 3.59$  \\
                                                  %& $5\%$  &  $49.16 \pm 0.30$  &  $46.65 \pm 0.83$  &  $\vsecondSig{54.69} \pm 0.33$  &  $\vfirstSig{56.29} \pm 0.42$  &  $38.00 \pm 0.37$  &  $39.73 \pm 0.37$  &  $43.24 \pm 1.81$  \\
                                                  %& $7\%$  &  $52.35 \pm 0.21$  &  $50.27 \pm 0.55$  &  $\vsecondSig{56.77} \pm 0.33$  &  $\vfirstSig{58.40} \pm 0.46$  &  $43.91 \pm 0.43$  &  $41.93 \pm 0.28$  &  $46.05 \pm 1.17$  \\
                                                  & $9\%$  &  $54.85 \pm 0.21$  &  $53.23 \pm 0.62$  &  $\vsecondSig{58.43} \pm 0.34$  &  $\vfirstSig{60.21} \pm 0.48$  &  $48.24 \pm 0.38$  &  $43.95 \pm 0.54$  &  $49.94 \pm 1.50$  \\
                                                  & $15\%$ &  $59.29 \pm 0.15$  &  $57.76 \pm 0.42$  &  $\vsecondSig{61.41} \pm 0.19$  &  $\vfirstSig{62.69} \pm 0.34$  &  $56.88 \pm 0.56$  &  $48.61 \pm 0.88$  &  $54.56 \pm 1.27$  \\
                                                  & $20\%$ &  $61.45 \pm 0.13$  &  $60.06 \pm 0.31$  &  $\vsecondSig{63.14} \pm 0.18$  &  $\vfirstSig{64.13} \pm 0.19$  &  $61.49 \pm 0.29$  &  $51.43 \pm 0.75$  &  $56.96 \pm 1.49$  \\
                                                  & $25\%$ &  $62.95 \pm 0.18$  &  $61.93 \pm 0.29$  &  $\vsecondSig{64.47} \pm 0.29$  &  $\vfirstSig{65.22} \pm 0.61$  &  $64.45 \pm 0.39$  &  $54.51 \pm 0.99$  &  $58.73 \pm 1.65$  \\
                                                  & time (ms)   &              32 &              7.1 &                            1226 &                           2341 &             116838 &             121530 &              129 \\
\multirow{7}{*}{\rotatebox[origin=c]{90}{\kiw{}}} 
                                                  & $3\%$  &               $26.02 \pm 0.27$  &  $\vsecondSig{26.23} \pm 0.46$  &   $\vfirstSig{33.92} \pm 0.50$  &  $23.59 \pm 0.43$  &              $20.02 \pm 0.68$  &  $20.13 \pm 0.56$  &   $1.11 \pm 0.26$  \\
                                                  %& $5\%$  &               $30.76 \pm 0.18$  &  $\vsecondSig{30.95} \pm 0.29$  &   $\vfirstSig{36.46} \pm 0.53$  &  $26.60 \pm 0.53$  &              $26.33 \pm 0.77$  &  $23.09 \pm 0.47$  &   $3.35 \pm 0.58$  \\

>>>>

\section{Experimental Analysis}\label{s:exp}

We now evaluate our edge sign classification methods on representative real-world datasets of
varying density and label regularity. This shows showing that our methods compete well against
existing approaches in terms of both predictive and computational performance. We are especially
interested in small training set regimes, and have restricted our comparison to the batch learning
scenario since all competing methods we are aware of have been developed in that setting only.

\paragraph{Datasets}

We considered five real-world classification datasets. The first three are directed \ssn{} widely
used as benchmarks for this task~(\eg{}\autocites{Leskovec2010}{shahriari2014ranking}{wu2016troll}):
In \wik{}, there is an edge from user $u$ to user $v$ if $v$ applies for an admin position and $u$
votes for or against that promotion. In \sla{}, a news sharing and commenting website, member $u$
can tag other members $v$ as friends or foes. Finally, in \epi{}, an online shopping website, user
$v$ reviews products and, based on these reviews, another user $u$ can display whether he considers
$v$ to be reliable or not. In addition to these three datasets, we considered two other \ssn{} where
the signs are inferred automatically, rather than given explicitly by the users.  In
\kiw{}~\cite{wikiedits11}, an edge from Wikipedia user $u$ to user $v$ indicates whether they edited
the same article in a constructive manner or not.\footnote{This is the
\href{http://konect.uni-koblenz.de/networks/wikisigned-k2}{KONECT version of the
\enquote{Wikisigned} dataset}, from which we removed self-loops.} Finally, in the
\aut{}~\cite{kumar2016structure} network, an author $u$ cites another author $v$ by either endorsing
or criticizing $v$'s work. The edge sign is derived by classifying the citation sentiment with a
simple, yet powerful, keyword-based technique using a list of positive and negative words. See
\cite{kumar2016structure} for more details.\footnote{We again removed self-loops and merged
multi-edges which are all of the same sign.}

\begin{table*}[bt]
  \centering
  \caption{Dataset properties. The 5th column gives the fraction of positive labels. The last two
  columns provide two different measures of label regularity ---see main text.\label{tab:dataset}}
  \begin{tabular}{lrrrrrrrr}
    \toprule
    Dataset &       $|V|$ &       $|E|$ &$\frac{|E|}{|V|}$ & $\frac{|E^+|}{|E|}$ & $\frac{\Psi^2_{G''}(Y)}{|E|}$ & $\frac{\Psi_G(Y)}{|E|}$ \\ % & twin edges & twin disagreement \\
    \midrule
    \aut{} &   \np{4831} &  \np{39452} & 8.1  &              72.3\% &                    .076 &                       .191 \\ %&      5.07\% &            27.10\% \\
    \wik{} &   \np{7114} & \np{103108} & 14.5 &              78.8\% &                    .063 &                       .142 \\ %&      5.56\% &            10.04\% \\
    \sla{} &  \np{82140} & \np{549202} & 6.7  &              77.4\% &                    .059 &                       .143 \\ %&     17.74\% &             4.00\% \\
    \kiw{} & \np{138587} & \np{740106} & 5.3  &              87.9\% &                    .034 &                       .086 \\ %&      6.55\% &             2.09\% \\
    \epi{} & \np{131580} & \np{840799} & 6.4  &              85.3\% &                    .031 &                       .074 \\ %&     30.83\% &            14.64\% \\
    \bottomrule
  \end{tabular}
\end{table*}

\autoref{tab:dataset} summarizes statistics for these datasets. We note that most edge labels are
positive. Hence, test set accuracy is not an appropriate measure of prediction performance. We
instead evaluated our performance using the so-called Matthews Correlation Coefficient
(MCC)~\autocite{MCC00}), defined as \Todo{For completeness, we could also report accuracy results in
the supplementary material.}  
\[
	\mathrm{MCC} = \frac{tp\times tn-fp\times fn}%
        {\sqrt{ (tp + fp) ( tp + fn ) ( tn + fp ) ( tn + fn ) } } = \pm \sqrt{\frac{\chi^2}{n}}
\]
MCC combines all the four quantities found in a binary confusion matrix ($t$rue $p$ositive, $t$rue
$n$egative, $f$alse $p$ositive and $f$alse $n$egative) into a single metric which ranges from $-1$
(when all predictions are incorrect) to $+1$ (when all predictions are correct) through $0$ (when
predictions are made uniformly at random).

Although the semantics of the edge signs is not the same across these networks, we can see from
\autoref{tab:dataset} that our generative model essentially fits all of them. Specifically, the last
two columns of the table report the rate of label (ir)regularity, as measured by
$\Psi^2_{G''}(Y)/|E|$ (second-last column) and $\Psi_{G}(Y)/|E|$ (last column), where 
\[
\Psi^2_{G''}(Y) = \min_{(\bp,\bq)} \left(f_{E_0}(\bp,\bq) + f_{E\setminus E_0}(\bp,\bq)\right)\,,
\]
$f_{E_0}$ and $f_{E\setminus E_0}$ being the quadratic criterions of \autoref{ss:passive}, viewed as
functions of both $(\bp,\bq)$, and $y_{i,j}$ when all labels are known, and $\Psi_{G}(Y)$ is the
label regularity measure adopted in the online setting, as defined in \autoref{s:prel}. It is
reasonable to expect that higher label irregularity corresponds to lower prediction performance.
This trend is in fact confirmed by our experimental findings: whereas \epi{} tends to be easy,
\aut{} tends to be hard, and this holds for all algorithms we tested, even if they do not explicitly
comply with our inductive bias principles. Moreover, $\Psi^2_{G''}(Y)/|E|$ tends to be proportional
to $\Psi_{G}(Y)/|E|$ across datasets, hence confirming the anticipated connection between the two
regularity measures.

@@

\paragraph{Algorithms and parameter tuning.} 
We compared the following algorithms:

\begin{enumerate}[label=\textbf{\arabic*.}]
  \item The label propagation algorithm of \autoref{ss:passive} (referred to as \uslpropGsec{}).
    The actual binarizing threshold was set by cross-validation on the training set.

  \item The algorithm analyzed at the beginning of \autoref{s:algbatch}, which we call \usrule{}
    (Bayes Learning Classifier based on \emph{tr}ollness and \emph{un}trustworthiness). After
    computing $\htr(u)$ and $\hun(u)$ on training set $E_0$ for all $u \in V$ (or setting those
    values to $\frac{1}{2}$ in case there is no outgoing or incoming edges for some node), we use
    equation \eqref{eq:predictor} and estimate $\tau$ on $E_0$.

  \item A logistic regression model where each edge $(u,v)$ is associated with the features
    $[1-\htr(u), 1-\hun(v)]$ computed again on \trainset{} (we call this method \uslogregp{}). Best
    binary thresholding is again computed on \trainset{}. Experimenting with this logistic model
    serves to support the claim we made in the introduction that our generative model in
    \autoref{s:gen} is a good fit for the data.

  \item  The solution obtained by directly solving the unregularized problem \eqref{e:quadratic}
    through a fast constrained minimization algorithm (referred to as \qoptim{}). Again, the actual
    binarizing threshold was set by cross-validation on the training set.\footnote{We have also
    tried to minimize \eqref{e:quadratic} by removing the $[-1,+1]$ constraints, but got similar MCC
    results as the ones we report for \qoptim{}}

  \item  The matrix completion method from~\autocite{LowRankCompletion14} based on \complowrank{}
    matrix factorization. Since the authors showed their method to be robust to the choice of the
    rank parameter $k$, we picked $k=7$ in our experiments.

  \item A logistic regression model built on \comptriads{} features derived from status
    theory~\autocite{Leskovec2010}.

  \item The PageRank-inspired algorithm from \autocite{wu2016troll}, where a recursive notion of
    trollness is computed by solving a suitable set of nonlinear equations through an iterative
    method, and then used to assign ranking scores to nodes, from which (un)trustworthiness features
    are finally extracted for each edge. We call this method \compranknodes{}. As for hyperparameter
    tuning ($\beta$ and $\lambda_1$ in~\autocite{wu2016troll}), we closely followed the authors'
    suggestion of doing cross validation.

  \item  The last competitor is the logistic regression model whose features have been build
    according to \autocite{Bayesian15}. We call this method \compbayesian{}.
\end{enumerate}

The above methods can be roughly divided into \emph{local} and \emph{global} methods. A local method
hinges on building local predictive features, based on neighborhoods: \usrule{}, \uslogregp{},
\comptriads{}, and \compbayesian{} essentially fall into this category. The remaining methods
(\uslpropGsec{}, \complowrank{}, and \compranknodes{}) are global in that their features are
designed to depend on global properties of the graph topology.

@@

Looking at undirected graphs with a global perspective, \textcite{LowRankCompletion14} consider the
observed adjacency matrix $A$, made of the edges in \trainset{}, as a noisy sampling of the
adjacency matrix $A^\star$ of an underlying complete graph satisfying the weakly balance condition
(that is with no triangle containing only one negative edge).
%TODO the previous sentence is too long
This condition implies the existence of a small number $k$ of node clusters with positive links
within clusters and negative links across clusters, which in turn implies $\rank A^\star=k$.  By
recovering a complete matrix $\tilde{A}$ which matches the non-zeros entries of $A$, it is possible
to predict the sign of $\eij \notin \trainset$ as $\yhat_{i,j} = \sgn(\tilde{A}_{i,j})$. Although
the exact version of this problem is NP-hard, authors assume that $k$ is an hyperparameter known
beforehand\footnote{The method is showed to be robust to the choice of its value, thus we pick
$k=7$.} and look for two matrices $W,H\in \mathbb{R}^{k\times|V|}$ which minimise a sigmoidal
reconstruction loss with $A$, subject to a regularization term. We refer to this method as
\emph{\complowrank{}}.

The status theory heuristic posits that a positive link from $i$ to $j$ denotes that user $i$
considers user $j$ as having a higher status or skill than himself~\autocite{Leskovec2010}. It
implies that among the 16 possible triads of three nodes with two directed signed edges among them,
some must be more represented than others. The \emph{\comptriads{}} method exploits this fact by
counting for each edge in the training set how frequently it is involved in each of the 16 triad
types. Based on these edge features and 7 degree features of each edge endpoints, a Logistic
Regression model is trained and used to predict the sign of the test edges.

Inspired by the PageRank algorithm, \textcite{wu2016troll} propose a recursive definition of
trollness based on the opinion of one node's neighbors weighted by their own trollness, which allow
to assign a trustworthiness $\pi(i)$ to each node $i$ through a set of non linear equations solved
by an iterative method. This $\pi$ values are used to rank the nodes by computing their Reputation
and Optimism scores, thus providing four features for each edge, which are in turn used to train a
Logistic Regression model for the classification task. We refer to this method as
\emph{\compranknodes{}}.  Computing $\pi$ requires to choose two hyperparameters of the model,
$\beta$ and $\lambda_1$ and the authors suggest to do it by holding out some training data to
perform cross validation but in practice we found little difference in performance in our
implementation while it consumes a lot of time.

The last competitors is also local. \textcite{Bayesian15} note that  a node can belong to one of the
16 types based of whether is number of its positive (resp.  negative) outgoing (resp. incoming)
edges is zero or not.  The number of unobserved incoming and outgoing edges of each node $i$ let us
define its 16-dimensional vector $V_i$ containing the probability of transitioning to any other
state.  Then for each edge \eij we compute the outer product of $V_i$ with $V_j$ and also include
additional features such as triads count and degree information before training a Logistic
Regression model. We refer to this method as \emph{\compbayesian{}}.
\fi

@@

  \caption{MCC with increasing training set size, with one standard deviation over 12 random
  sampling of $\trainset$. The last four columns refer to the methods we took from the literature.
For the sake of readability, we multiplied all MCC values by 100. The best number in each row is
highlighted in \textbf{\textcolor{brown}{bold brown}} and the second one in
\textit{\textcolor{red}{italic red}}. If the difference is statistically significant ($p$-value of a
paired Student's $t$-test less than $0.005$), the best score is underlined. The ``time" rows contain
the time taken to train on a $15\%$ training set.\label{tab:all_mcc}}
